### 2018-09-19 Manuscripts/cog-psych/revision/cogpsych.tex
Incorporated minor updates suggested by Ken (in particular the unidimensional sort bias bit.)

To successfully generate something novel, what is generated must be different
from what is already known. This fundamental constraint, ``being different'', or
contrasting from other categories in the relevant domain, is the focus of our
work. Intuitively, the influence of contrast may be expected in a number of
real-life scenarios involving the generation of new categories. For instance, a
college student who has been eating instant noodles for a week may be more
inclined to seek out something healthier and less carbohydrate-heavy (e.g., a
Caesar salad) compared to something similar (a different flavour of instant
noodles). An actor who is dissatisfied with being typecast as a villain may seek
out roles as heroes as opposed to playing other antagonistic characters. A more
compelling example is seen in recent evidence by \cite{askin2017} who found that
new songs that were more different to their peers were generally more popular
than new songs that were similar to others. It appears that once there is the
motivation to generate something new, there is an associated motivation to
generate something \emph{different}.

@@

Interestingly, the unidimensional sort bias that is commonly observed in
unsupervised categorization
\citep{imai1965discriminability,milton2004influence,ahn1992two}, where
participants sort unlabelled family resemblance category exemplars by focusing
on a single feature, was not consistently observed in our tasks. In our
experiments we only observed this bias in the row condition of Experiment 1,
whereas in other conditions participants generated categories that typically
varied along both dimensions. This may be unsurprising at face value, since
participants in conditions other than the row condition were trained on a prior
category that varied on both dimensions. However, these results suggest that the
unidimensional sort bias is not universal to all types of category construction
-- or more specifically, the unidimensional bias may not apply when participants
need to generate categories with entirely new exemplars.

>>>>

To successfully generate something novel, what is generated must be different
from what is already known. This fundamental constraint, ``being different'', or
contrasting from other categories in the relevant domain, is the focus of our
work. Intuitively, the influence of contrast may be expected in a number of
real-life scenarios involving the generation of new categories. For instance, a
college student who has been eating instant noodles for a week may be more
inclined to seek out something healthier and less carbohydrate-heavy (e.g., a
Caesar salad) compared to something similar (a different flavour of instant
noodles). An actor who is dissatisfied with being typecast as a villain may seek
out roles as heroes as opposed to playing other antagonistic characters. An
especially compelling example can be seen in \cite{askin2017} who found that
new songs that were more different to their peers were generally more popular
than new songs that were similar to others. It appears that once there is the
motivation to generate something new, there is an associated motivation to
generate something \emph{different}.

@@

Interestingly, the unidimensional sort bias that is commonly observed in
unsupervised categorization
\citep{imai1965discriminability,milton2004influence,ahn1992two}, where
participants sort unlabelled family resemblance category exemplars by focusing
on a single feature, was not consistently observed in our tasks. In our
experiments we only observed this bias in the row condition of Experiment 1,
whereas in other conditions participants generated categories that typically
varied along both dimensions. This may be unsurprising at face value, since
participants in conditions other than the row condition were trained on a prior
category that varied on both dimensions. However, these results suggest that the
unidimensional sort bias is not universal to all types of category construction
-- or more specifically, the unidimensional bias may not apply when participants
need to generate categories with entirely new exemplars. Additionally, the
unidimensional sort bias is observed in category construction tasks in which
participants develop a category-level organization of a provided domain while
category generation tasks involve generating a new category relative to one that
was just acquired (one might say that participants are forging a domain from a
category rather than categories from a domain). Since the more creative task of
generating a category from scratch (with respect to an established category)
does not invoke the unidimensional bias, this suggests that a more creatively
demanding version of a sort task might also reduce the bias.
